MLH1 Results Outline

### 2020-02-21 doc/Results.rmd
short notes on the IFD part

DO THE HIGH QUALITY

@@

Q1.IFD.raw_BIV.short.ttest_KAZ <-  t.test( (real.short.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
                ( real.short.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)

#THIS BIV OBS are also SIG

@@

@@

nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
nrm.IFD_KAZ_BIV.data

@@

nrm.IFD_PWD_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
nrm.IFD_PWD_BIV.data
#this might

nrm.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male GMSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
nrm.IFD_MSM_BIV.data

nrm.IFD_SKIVE_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="SKIVE" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male SKIVE")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
nrm.IFD_SKIVE_BIV.data

nrm.IFD_G_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
nrm.IFD_G_BIV.data
```

The plots above are focusing on the male pattern -- first meant to look at the pattern in KAZ -- WHICH DOESN"T have significant sex difference in NRM IFD -- INSTEAD it has a low rate of 2CO bivalents with very short IFD(nrm)  --- what if these were mistakes? they were supposed to be 1CO -- but somehow a second CO formed anyways .. and very close by

@@

whats the expectations following the confonding proportions? 
high rec strains have more 2COs so they there is potential for more variance // chromsome specific effects   where as for the low rec strains -- which have fewer 2COs, there might (big might) be less chromosome specific effects.

>>>>

DO THE HIGH QUALITY dataset tests

@@

Q1.IFD.raw_BIV.short.ttest_KAZ <-  t.test( (real.short.bivData %>% filter(sex == "male") %>% filter(strain =="KAZ" )  %>% filter(hand.foci.count == "2") )$IFD1_PER,
                ( real.short.bivData %>% filter(sex == "female") %>% filter(strain =="KAZ") %>%  filter(hand.foci.count == "2"))$IFD1_PER)
```

```{r Q1.show.IFD.things, echo=FALSE}

@@

nrm.IFD_PWD.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
nrm.IFD_PWD.female_BIV.data

nrm.IFD_LEW.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="LEW" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female LEW")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
nrm.IFD_LEW.female_BIV.data

nrm.IFD_G.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
nrm.IFD_G.female_BIV.data

nrm.IFD_WSB.female_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="WSB" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "female")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, female WSB")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))
nrm.IFD_WSB.female_BIV.data    #18nov17_WSB_f2 has continuous ifd range

@@

nrm.IFD_KAZ_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="KAZ" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male KAZ")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))

@@

nrm.IFD_PWD_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="PWD" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male") ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male PWD")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))

#this might

nrm.IFD_MSM_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="MSM" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male GMSM")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))

nrm.IFD_SKIVE_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="SKIVE" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male SKIVE")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))

nrm.IFD_G_BIV.data <- ggplot( (Curated_BivData %>% filter(strain =="G" ) %>%  filter(hand.foci.count == "2") %>%  filter(sex == "male")  ) , aes(y = IFD1_PER, x= fileName, color=chromosomeLength))+geom_point()+ggtitle("Normalized Interfocal Distances, male G")+facet_wrap(~mouse, scales = "free")+ylim(c(0,1))

see the reference plots within this dir, --

this section -- all strains but KAZ, males have a higher mean normalized IFDs

the female norm IFD are centered at ~50% -- but they also seem to have a higher rate of short (<25%) normalized IFDs

The KAZ strain is the only strain without significant p value for the normalized IFDs, due to KAZ males having a higher rate (enrichment of short normalized IFDs.

There seems to be a 25% norm.IFD cutoff (in females) -- but some mice seem to have less of this cut
off.
Is there a default expectation for why .25 would be the cut off...(given female uniform CO positioning?

<The plots above are focusing on the male pattern -- first meant to look at the pattern in KAZ -- WHICH DOESN"T have significant sex difference in NRM IFD -- INSTEAD it has a low rate of 2CO bivalents with very short IFD(nrm)  --- what if these were mistakes? they were supposed to be 1CO -- but somehow a second CO formed anyways .. and very close by

@@

whats the expectations following the confonding proportions? 
high rec strains have more 2COs so they there is potential for more variance // chromsome specific effects   where as for the low rec strains -- which have fewer 2COs, there might (big might) be less chromosome specific effects.>
